Lab 9: Multinomial and Ordered Probit Models: Answers

### 2023-03-08 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit.Rmd
Add files via upload

---
title: "Lab 9: Multinomial and Ordered Probit Models"
author: "Eric Parajon"
output: pdf_document
---

@@

# Lab Assignment

@@

Generate a cumulative predicted probability plot.

@@

>>>>

---
title: 'Lab 9: Multinomial and Ordered Probit Models'
author: "Eric Parajon"
output:
  html_document:
    df_print: paged
---

@@

# Exercise

@@

Generate a cumulative predicted probability plot varying the prestige of the respondent's occupation.

@@

#Cleaning the data
hw<-hw %>% 
   mutate(warm = case_when(____)) %>%
    mutate(warm = factor(warm, levels=c("Strongly disagree","Disagree","Agree","Strongly agree")))

#Create the model
warm_model <- polr(____, Hess = TRUE)
summary(warm_model)

#What is the effect of increasing each predictor by one unit?
#How do the odds of moving from a lower to adjacent higher category change?
coefs <- warm_model$_____
#odds ratios
_____(coefs)

#Check the proportional odds assumption for a variable

#Generate a cumulative predicted probability plot.
table(hw$male); table(hw$white); mean(hw$ed)

hw_2 <- expand.grid(_____,
                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))

probs <- predict(____, type = "probs") #type = is different than for binomial model

cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
#transpose into columns for plotting and merging with values of gpa

plot_data <- data.frame(hw$prst, cumul_probs)

#Generate a cumulative predicted probability plot.
ggplot(data=plot_data, aes(x=hw.prst))+
      ______

### 2023-03-08 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit_Answers.Rmd
Add files via upload

---
title: "Lab 9: Multinomial and Ordered Probit Models"
author: "Eric Parajon"
output: pdf_document
---

@@

Load necessary packages:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
rm(list=ls())

@@

```{r, eval = F}
hw <- read.csv(url("http://faculty.washington.edu/cadolph/mle/ordwarm2.csv"))

>>>>

---
title: 'Lab 9: Multinomial and Ordered Probit Models: Answers'
author: "Eric Parajon"
output:
  html_document:
    df_print: paged
---

@@

Load necessary packages:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messsage = FALSE, warning=FALSE)
rm(list=ls())

@@

```{r}
hw <- read.csv(url("http://faculty.washington.edu/cadolph/mle/ordwarm2.csv"))

#Cleaning

hw<-hw %>% 
   mutate(warm = case_when(.$warm == 1 ~ "Strongly disagree", 
                          .$warm==2 ~ "Disagree", 
                          .$warm==3 ~ "Agree", 
                          .$warm==4 ~ "Strongly agree")) %>%
    mutate(warm = factor(warm, levels=c("Strongly disagree","Disagree","Agree","Strongly agree")))

#Create the model
warm_model <- polr(warm~male+white+age+ed+prst, 
           data = hw, method = "probit", Hess = TRUE)
summary(warm_model)

#What is the effect of increasing each predictor by one unit?
#How do the odds of moving from a lower to adjacent higher category change?
coefs <- warm_model$coefficients[1:5]
#odds ratios
exp(coefs)

#Check the proportional odds assumption for male
m1_prop_1 <- glm(I(as.numeric(warm) >= 2) ~ male, family = "binomial",
                 data = hw)

m1_prop_2 <- glm(I(as.numeric(warm) >= 3) ~ male, family = "binomial",
                 data = hw)

#male=0
check_1 <- m1_prop_1$coef[1] - m1_prop_2$coef[1]
check_1

#male=1
check_2 <- (m1_prop_1$coef[1] + m1_prop_1$coef[2]) - 
  (m1_prop_2$coef[1] + m1_prop_2$coef[2])
check_2

#Generate a cumulative predicted probability plot.
table(hw$male); table(hw$white); mean(hw$ed)
hw_2 <- expand.grid(male = 1, #modal value
                      white = 1, #modal value
                      ed=mean(hw$ed),
                      age=mean(hw$age),
                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))

probs <- predict(warm_model, hw_2, type = "probs") #type = is different than for binomial model
head(rowSums(probs)) #just to double check
cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
#transpose into columns for plotting and merging with values of gpa
head(cumul_probs)
plot_data <- data.frame(hw$prst, cumul_probs)

#plot data
ggplot(data=plot_data, aes(x=hw.prst))+
    geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.disagree), se=FALSE)+
    geom_smooth(data=plot_data, aes(x=hw.prst, y=Disagree), se=FALSE)+
    geom_smooth(data=plot_data, aes(x=hw.prst, y=Agree), se=FALSE)+
      geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.agree), se=FALSE)+
    geom_ribbon(data=plot_data, aes(ymin=0, ymax=Strongly.disagree, fill="Strongly.disagree"))+
    geom_ribbon(data=plot_data, aes(ymin=Strongly.disagree, ymax=Disagree,
                                    fill="Disagree"))+
    geom_ribbon(data=plot_data, aes(ymin=Disagree, ymax=Agree,
                                    fill="Agree"))+
         geom_ribbon(data=plot_data, aes(ymin=Agree, ymax=Strongly.agree,
                                    fill="Strongly.agree"))+
    labs(x="Values of prestige of occupation", y="Cumulative Predicted Probability")+
    scale_fill_discrete(name="Hard Work")+
    theme_bw()
